---
title: "visualization_2"
author: "Ziqing Wang"
date: "2022-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In addition to figure sizing, I include a few other figure preferences in global options declared at the outset of each .Rmd file (this code chunk just gets copy-and-pasted to the beginning of every new file).
```{r}
library(tidyverse)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```
There are ways to set color preferences globally as well (for example, to use viridis color palettes everywhere, all the time, without setting these options in your RMarkdown file), although they’re a bit more involved.


## Load the NOAA data again
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## label and scale
As a starting point, let’s revisit the scatterplot of tmax against tmin made in Visualization Pt 1, but make some improvements.
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs( # add labels
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) +
  scale_x_continuous( # cheange scale for x axis
    breaks = c(-15, 0, 15), # change places to put x "ticks" to different from default
    labels = c("-15º C", "0", "15º C")) + # assign new labels to these "ticks"
  #scale_color_hue(h = c(10, 30), (not recommended)
  #                name="location") # change color hue/scheme
  scale_y_continuous( # can also transform data on a certain axis using scale
    trans = "sqrt", 
    position = "right") +
  viridis::scale_color_viridis( # better color palette; useful for continuous color palette
    name = "location",
    discrete = T
  )

```

## Theme
Be careful with the orders when adding multiple themes!
```{r}
ggp_temp_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") # a theme where the legend position is at the bottom of the graph
  

ggp_temp_plot
```

## Data argument in geom_*
We’ve seen that where an aesthetic gets mapped to a variable matters – setting aes(color = name) in ggplot can yield different results than the same setting in geom_point(). This arises from the way that ggplot objects inherit aesthetic mappings, and it turns out there’s a similar thing with the data used to make a plot.

In a contrived example, we can split weather_df into separate datasets for Central Park and Waikiki. Then we use one in the ggplot() call and another in geom_line():

```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park) # overlay another set of data using anothe geometry (in this case, line instead of points)
```

## Patchwork
We’ve seen facetting as an approach to create the “same plot” for several levels of a categorical variable, and this can get you pretty far. Sometimes, though, you want to show two or three fundamentally different plots in the same graphic: you may want to juxtapose a scatterplot and a boxplot, or show scatterplots illustrating relationships between different variables. In this case, a solution is to create each of the panels you want separately and combine panels using tools in the patchwork package:

```{r}
tmax_tmin_p = 
  weather_df %>% 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")


(tmax_tmin_p + prcp_dens_p) / tmax_date_p

```

In patchwork,
+: put side by side
/: put up and down

## data manipulation
Often, struggles with ggplot are struggles with data tidying in disguise. Viewing data manipulation as part of the visualization process will often be your path to success! Put differently, the behavior of your plot depends on the data you’ve supplied; in some cases, it’s easier to control behavior through data manipulation than it is through the plot code.

This is particularly true for the order of categorical or factor variables in plots. Categorical variables will be ordered alphabetically; factors will follow the specified order level that underlies the variable labels. You can change the order level of a factor variable to your specified preference using forcats::fct_relevel or according to the value of another variable using forcats::fct_reorder.

A first example reorders name “by hand”:
```{r}
weather_df %>%
  # explicitly make name a factor variable and manually assign the level order (instead of using the default alphabetical order)
  mutate(name = forcats::fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")

```

A second example reorders name according to tmax values in each name:
```{r}
weather_df %>%
  mutate(name = forcats::fct_reorder(name, tmax)) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")

```

Our emphasis on data tidiness in previous examples is helpful in visualization. The code below imports and tidies the PULSE data, and creates a plot showing BDI score across visits. Some steps that are helpful in retrospect are using pivot_longer to organize the BDI score and visit time variables, and organizing the visit time variable into a factor with an informative ordering.

```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  select(id, visit, everything()) %>%
  mutate(
    visit = recode(visit, "bl" = "00m"),
    visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)

ggplot(pulse_data, aes(x = visit, y = bdi)) + 
  geom_boxplot()

```













